---
title: "Multiple Linear Regression"
output:
  pdf_document: default
---

Here, we will use the MTS package to compute forecasts of the log gas consumption, basing ourselves on : 
- the precedent values of the log gas consumption 
- the selected temperature metrics, that is to say min_temp_val and heat_days_val
- the computed dummy variables extreme_heat and extreme_cold 

### Load data
As usual, we load the datasets. 

```{r}
library(readr)
library(forecast)
gas <- na.omit(read_csv("data/gas.csv", col_types = cols(date = col_date(format = "%Y-%m-%d"))))
temp <- read_csv("data/temp.csv", col_types = cols(date = col_date(format = "%Y-%m-%d")))
google <- read_csv("data/google.csv", col_types = cols(date = col_date(format = "%Y-%m-%d")))
```

As said before, our datasets span over different periods of time: while gas consumption, our target value goes from January 1973 to July 2016, temperature data start in January 1895 and end in February 2017. Finally, data from Google Trends go from january 2004 and up to march 2017.

We will thus need to use the predicted values : 
- of min_temp_val and heat_days_val from February 2017 to July 2017
- of extreme_heat and extreme_cold from March 2017 to July 2017. 

Loading these predictions obtained previously: 
```{r}
library(readxl)
inter_preds <- read_excel("data/inter_preds.xlsx")
inter_preds
```
Here extreme_heat and extreme_cold have NAs for april because we have the actual values in the "google" dataset. 

### Preparation

Let us merge the dataframes
```{r}
gas_temp = merge(gas, temp, by = "date", all.x = TRUE)
gas_goo = merge(gas, google, by = "date", all = FALSE)
all = merge(gas_temp, google, by= "date",  all = FALSE)

extreme <- all[,c(16,17,18)]

# extreme_heat will be 1 when heatwave > mean(heatwave)
extreme_heat <- as.numeric(extreme$heatwave > mean(extreme$heatwave))
# extreme_cold will be 1 when snow_storm > mean(snow_storm) and extreme_weather > mean(extreme_weather)
extreme_cold <- as.numeric(extreme$snow_storm > mean(extreme$snow_storm) & extreme$extreme_weather > mean(extreme$extreme_weather))

all = cbind(all, extreme_heat, extreme_cold)
```

Let us create the test and train set
```{r}
all_test = all[all$date >= '2015-06-15' & all$date < '2016-06-15',]
all_train = all[all$date < '2015-06-15',]
nrow(all_train)
```

### Multiple linear regression 

We will test different combinasons. 

```{r}
train = all_train[,c("log_gas_cons", "min_temp_val", "heat_days_val", "extreme_heat", "extreme_cold")]
test = all_test[,c("log_gas_cons", "min_temp_val", "heat_days_val", "extreme_heat", "extreme_cold")]

linreg_all = lm(log_gas_cons ~., data = train)
pred_linreg = predict.lm(linreg_all, newdata = test)
accuracy(test$log_gas_cons, pred_linreg)

summary(linreg_all)

train = all_train[,c("log_gas_cons", "min_temp_val", "heat_days_val")]
test = all_test[,c("log_gas_cons", "min_temp_val", "heat_days_val")]

linreg_2 = lm(log_gas_cons ~., data = train)
pred_linreg = predict.lm(linreg_2, newdata = test)
accuracy(test$log_gas_cons, pred_linreg)

summary(linreg_2)

train = all_train[,c("log_gas_cons", "min_temp_val", "heat_days_val","extreme_heat")]
test = all_test[,c("log_gas_cons", "min_temp_val", "heat_days_val", "extreme_heat")]

linreg_3 = lm(log_gas_cons ~., data = train)
pred_linreg = predict.lm(linreg_3, newdata = test)
accuracy(test$log_gas_cons, pred_linreg)

summary(linreg_3)
```

The best linear regression model is thus obtained with "log_gas_cons", "min_temp_val", "heat_days_val" and "extreme_heat". 

Let's convert in Time series
```{r}
train = all_train[,c("log_gas_cons", "min_temp_val", "heat_days_val","extreme_heat")]
test = all_test[,c("log_gas_cons", "min_temp_val", "heat_days_val", "extreme_heat")]

strain = ts(train, start = c(2004, 1), frequency=12)
train = window(strain, start = c(2004, 1), end = c(2015, 2))
stest = ts(test, start = c(2015,3), frequency=12)
test = window(stest, start = c(2015,3), end = c(2016, 2))
```

### Rolling Window Evaluation
To try more model, we would like to setup a rolling window. Let us find the optimal size of that window using the multiple linear regression we just selected. 

```{r}
library(ggplot2)
err = data.frame("RMSE" = rep(0, 2014-2004+1), row.names = seq(2004, 2014))
for (y in 2004:2014){
    model = lm(log_gas_cons ~., data = window(train, start=c(y,1)))
    rmse = accuracy(predict.lm(model, newdata = test),test[,1])[2]
    err[as.character(y),] = rmse
}
ggplot(err)+geom_line(aes(x=as.numeric(row.names(err)), y=RMSE))
```
So we'll take 2 years of data (since the result is 2012). 

#### Final prediction : including both the predicted (when missing) and the actual temperatures

Preparation of the databases (complicated here because we have a datapoint more in extreme_heat, ie google database):

```{r}
true_extreme_heat <- ts(as.numeric(google$heatwave > mean(google$heatwave)), start = c(2004, 1), frequency=12)
pred_extreme_heat <- ts(inter_preds$extreme_heat[-1], start = c(2017, 4), frequency=12)

true_min_temp_val = ts(temp$min_temp_val, start = c(1895, 1), frequency=12)
true_heat_days_val = ts(temp$heat_days_val, start = c(1895, 1), frequency=12)

pred_min_temp_val = ts(inter_preds$min_temp_val, start = c(2017, 3), frequency=12)
pred_heat_days_val = ts(inter_preds$heat_days_val, start = c(2017, 3), frequency=12)

tslog_gas_cons = ts(gas$log_gas_cons, start = c(1973,1), frequency = 12)

data = as.data.frame(cbind(
    window(tslog_gas_cons, start = c(2004,1), end = c(2016, 6)), 
    window(true_min_temp_val, start = c(2004,1), end = c(2016, 6)), 
    as.double(window(true_heat_days_val, start = c(2004,1), end = c(2016, 6))), 
    as.double(window(true_extreme_heat, start = c(2004,1), end = c(2016, 6)))
    )) 
colnames(data) = c("log_gas_cons", "min_temp_val", "heat_days_val", "extreme_heat")

nd_extreme_heat = ts(c(window(true_extreme_heat, start = c(2016, 7)),pred_extreme_heat), start = c(2016, 7), frequency=12)
nd_min_temp_val = ts(c(window(true_min_temp_val, start = c(2016, 7)),pred_min_temp_val), start = c(2016, 7), frequency=12)
nd_heat_days_val = ts(c(window(true_heat_days_val, start = c(2016, 7)),as.double(pred_heat_days_val)), start = c(2016, 7), frequency=12)

newdata = as.data.frame(cbind(as.double(nd_min_temp_val), as.double(nd_heat_days_val), as.double(nd_extreme_heat)))
colnames(newdata) = c("min_temp_val", "heat_days_val", "extreme_heat")
```

Fitting the model on all the dataset and plotting the predictions : 

```{r}
model = lm(log_gas_cons ~., data = data)
prediction = predict.lm(model, newdata = newdata)
prediction = ts(prediction, start = c(2016,7), frequency = 12)
plot(prediction, main = "Prediction of Multiple linear regression on all the dataset", ylab = "log gas consumption")
plot(exp(prediction), main = "Prediction of Multiple linear regression on all the dataset", ylab = "actual gas prediction")
```
Fitting the model on only the two last years and plotting the predictions : 

```{r}
data2 = data[127:150,]
model2 = lm(log_gas_cons ~., data = data2)
prediction2 = predict.lm(model2, newdata = newdata)
prediction2 = ts(prediction2, start = c(2016,7), frequency = 12)
plot(prediction2, main = "Prediction of Multiple linear regression on 2 years of the dataset", ylab = "log gas consumption")
plot(exp(prediction2), main = "Prediction of Multiple linear regression on 2 years of the dataset", ylab = "actual gas prediction")
```