---
title: "Univariate Forecasts"
output: html_notebook
---

In this part we evaluate univariate forecasts of the gas consumption.

### Load the data
```{r}
library(readr)
gas <- na.omit(read_csv("data/gas.csv", col_types = cols(date = col_date(format = "%Y-%m-%d"))))
```

### Stationarity and integration
```{r}
library(forecast)
library(fpp)
tsdisplay(gas$log_gas_cons,main="Log transform of Gas Consumption")
tsdisplay(diff(gas$log_gas_cons, 3), main="Third Order Differientiation")
tsdisplay(diff(gas$log_gas_cons, 12), main="Twelvth Order Differientiation")
```
Unsurprisingly, we notice a strong seasonality, still observable after seasonal differencing. Exploring further the dynamics of the log consumption with annual differencing yields a time series that looks stationary.

### Models evaluation
Let us convert data to a time series
```{r}
lgas = ts(gas$log_gas_cons, start = c(1973, 1), frequency=12)
train = window(lgas, start = c(1973, 1), end = c(2015, 6))
test = window(lgas, start = c(2015,7), end = c(2016, 6))
```

We would like to evaluate models using a rolling window. Let us look at what R suggests:

## Time Series Decomposition

TODO SEBASTIEN PENAY

## Seasonal ARIMA

### Auto Arima
```{r}
autofit = auto.arima(train, seasonal=TRUE)
print(autofit)
autofk = forecast(autofit,h=12)
accuracy(autofk,test)
```

Let us now try with our own seasonal ARIMA. Based on the ACF PACF of the annual differentiation, we try with an  ARIMA(4,0,1)(0,1,2).

### Custom Seasonal Arima
```{r}
custfit = Arima(train, order=c(4,0,1), seasonal=c(0,1,2), lambda=0)
print(custfit)
custfk = forecast(custfit,h=12)
accuracy(custfk,test)
```

Although the statistical tests and criterion seem better here, the error on train and test sets are higher, so we keep the auto ARIMA.

### Rolling Window Evaluation
To try more model, we would like to setup a rolling window. Let us find the optimal size of that window using the Auto ARIMA we just fitted.



